Integer Programming Formulations for the Shared Multicast Tree Problem

### 2018-12-27 smt/sec6-strong.tex
MI: Remarks from minor revision incorporated - mostly typos and grammar.

Let $A(D)=A\cap(D\times D)$ be the set of ordered pairs of distinct destination nodes.
For a tree $T$ solving Problem \ref{def:problem}, define the variables $x\in\{0,1\}^{A\times A(D)}$, where
$x_{ij}^{st}=1$ iff $(i,j) \in A$ is an arc on the path from $s$ to $t$ in arborescence $T^s$.
% \newline\newline
% $x_{ij}^{st}=
% \begin{cases}
%     1 & \text{if $(i,j) \in A$ is an arc on the path from $s$ to $t$ in $T$, $(s,t)\in A(D)$},\\
%     0 & \text{otherwise}.
% \end{cases}$
% \newline\newline
Model $\mathcal{X}_2$ is strengthened by constraints on $x$, which yields model $\mathcal{X}_3$: 
\newline
\newline
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                     % 
%             MODEL X3                %
%             (4 idx vars)            %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{subequations}[resume]\label{eq:smt-x3}
\begin{align}
\notag\makebox[0pt][l]{$\displaystyle{} \min \sum\limits_{(i,j) \in A} p_{ij} \sum\limits_{s \in D} \pi^s_{ij} $}\\
\text{s.t.} \notag \\
(\ref{con:dd:arrowFromDest}) - (\ref{con:dd:vardim}) &\text{ and }(\ref{con:dd:yvar}) - (\ref{con:vi:sumYImpSumX}) \notag\\
\label{con:mf:flowNormal} \sum\limits_{\substack{ j\in V_i}}x^{st}_{ij}-\sum\limits_{\substack{j\in V_i }}x^{st}_{ji} & = 0 & (s,t)\in A(D), i\in V\setminus\{s,t\},\\
\label{con:mf:flowDest} \sum\limits_{\substack{j\in V_t}}x^{st}_{jt}    & = 1  & (s,t)\in A(D),\\
\label{con:mf:fcap} x^{st}_{ij} &\leq X^{s}_{ij}& (i,j)\in A, (s,t)\in A(D),\\
\label{con:mf:fsym} x^{st}_{ij} &=  x^{ts}_{ji}& (i,j)\in A, (s,t)\in A(D),\\ 
\label{con:vi:sumFImpSumY} \sum_{k\in W_{ij}}x^{st}_{ik} & \leq \sum_{k\in W_{ij}} \pi^{s}_{jk} & (i,j)\in A, (s,t)\in A(D),\\
\label{con:mf:xydim} x&\in\{0,1\}^{A\times A(D)}. 
\end{align}~
\end{subequations}

Variable $x_{ij}^{st}$ is also interpreted as the \emph{flow of commodity $(s,t)$} along arc $(i,j)$.
Correspondingly, (\ref{con:mf:flowNormal})--(\ref{con:mf:flowDest}) can be viewed as flow conservation constraints,
stating that there exists a path from $s$ to $t$ in $T^s$.
Next, constraints (\ref{con:mf:fcap}) express the obvious relation between $X$ and $x$, that $(i,j)$ is an arc on the path from $s$ to $t$ only if it exists in $T^s$.
The flow symmetry equations (\ref{con:mf:fsym}) state that the path from $s$ to $t$ contains $(i,j)$ iff the reverse path contains the reverse arc $(j,i)$.
Finally, constraints (\ref{con:vi:sumFImpSumY}) reflect that if the path from $s$ to $t$ intersects $(i,j)$,
then the most expensive arc leaving $i$ in $T^s$ has cost $p_{ij}$ or more.

@@

Similarly to the extension $\mathcal{X}_3$ of $\mathcal{X}_2$ by $x$-variables, the model $\mathcal{F}_2$ is extended by variables with four node indices.
Analogously to $A(D)$, let $E(D_0)=E\cap (D_0\times D_0)$ be the set of pairs of \emph{unordered} destination nodes other than $s_0$.
Define variables $f\in\{0,1\}^{A\times E(D)}$, where
$f_{ij}^{st}=f_{ji}^{ts}$ takes the value 1 iff the path in $T^0$ from $s_0$ to $s$ intersects with the path from $s_0$ to $t$ at arc $(i,j)$.
% \newline\newline
% $f^{st}_{ij}=f^{ts}_{ij}=
% \begin{cases}
%     1 & \text{if the paths in $T^0$ from $s_0$ to respectively $s$ and $t$ intersect} \\
%       & \text{at $(i,j)$,} \\
%     0 & \text{otherwise.}
% \end{cases}$
% \newline\newline
%
Variables $f_{ij}^{st}$ can also be regarded as the \emph{joint flow} from $s_0$ to $s$ and $t$ \citep{Polzin}.
In the cited work, they are successfully used to strengthen formulations for the \textsc{Minimum Steiner Tree} problem.
The new variables enable formulation of an extended model denoted $\mathcal{F}_3$:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                     %
%             MODEL F3                %
%                                     %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{subequations}[resume]
\begin{align}
\notag\min & \sum\limits_{(i,j) \in A} p_{ij} \sum\limits_{s \in D} \pi^s_{ij}\\ \notag
\text{s.t.}& && \\
\notag(\ref{con:pf1:flow})-(\ref{con:vi:sumYImpSumXTrans}) & ~\text{and} ~(\ref{con:vi:Y1}) \\
\label{con:pf2:flowHook} \sum\limits_{\substack{ j \in V_i }}\left(f^{st}_{ij}-f^{st}_{ji}\right) & \leq 
  \begin{cases}
    1, & i = s_0, \\
    0, & i\in V_0,
  \end{cases} \quad \{s,t\}\in E(D_0), \\
\label{con:pf2:startInSource}  f^{st}_{ij} & \leq F^s_{ij} \quad (i,j)\in A, \{s,t\}\in E(D_0),\\
\label{con:pf2:stopInDest}  f^{st}_{ij} & \leq F^t_{ij} \quad (i,j)\in A, \{s,t\}\in E(D_0),\\
\label{con:pf2:stronger}  F^{s}_{ij}+F^{t}_{ij}-f^{st}_{ij} &\leq g_{ij} \quad (i,j)\in A, \{s,t\}\in E(D_0),\\ 
\label{con:pf2:sumFImpSumY} \sum\limits_{k\in W_{ij}}\left(F^{t}_{ik} + F^{s}_{ki} - f^{st}_{ik} - f^{st}_{ki}\right) & \leq \sum\limits_{k\in W_{ij}}  \pi^{s}_{ik} \quad (i,j)\in A,\{s,t\}\in E(D_0),\\
\label{con:pf2:sumFImpSumY2} \sum\limits_{k\in W_{ij}}F^{t}_{ik} & \leq \sum\limits_{k\in W_{ij}}  \pi^{0}_{ik} \quad (i,j)\in A, t\in D_0,\\
\label{con:pf2:dim} f&\in\{0,1\}^{A\times E(D_0)}.
\end{align}~
\end{subequations}

By (\ref{con:pf2:flowHook}), it is ensured that the joint flow is non-increasing by increasing distance from the root $s_0$.
While constraints (\ref{con:pf2:startInSource})--(\ref{con:pf2:stopInDest}) ensure $f_{ij}^{st}=0$ if $F_{ij}^s=0$ or $F_{ij}^t=0$,
inequalities (\ref{con:pf2:stronger}) impose $f_{ij}^{st}=1$ if $F_{ij}^s=F_{ij}^t=1$.
By virtue of (\ref{con:pf2:stopInDest}), (\ref{con:pf2:stronger}) replace the weaker capacity constraints (\ref{con:pf1:xfrel}).
Finally, (\ref{con:pf2:sumFImpSumY})--(\ref{con:pf2:sumFImpSumY2}) are valid inequalities justified by:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                     %
%         Proposition  		      %
%   (transformation x_{ij}^{st})      %
%                                     %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{prop}\label{prop:transX}
Assume $(F,f,g,\pi)$ satisfies (\ref{con:pf1:flow})--(\ref{con:vi:sumYImpSumXTrans}), (\ref{con:vi:Y1}), (\ref{con:pf2:flowHook})--(\ref{con:pf2:stronger}), and (\ref{con:pf2:dim}).
Let $T^0\subseteq G_g$ be the arborescence spanning $D$ (see Lemma \ref{lem:modelcorrect}),
and let $T^s$ be the arborescence obtained by redirecting the arcs in $T^0$ such that they point away from $s\in D$. Then, for all nodes $i$ and $j$ ($i\neq j$) spanned by $T^0$,
$$
F^t_{ij}+F^s_{ji}-f^{st}_{ij}-f^{st}_{ji} = 
	\begin{cases}
		1, ~\text{if $(i,j)$ is an arc on the path from~$s$ to $t$ in $T^s$,} \\
		0, ~\text{otherwise.}
	\end{cases}
$$
\end{prop}
%
\begin{proof}
Because $g_{ij}+g_{ji}\leq 1$, we have $F^t_{ij}+F^s_{ji}\leq 1$.
Constraints (\ref{con:pf2:startInSource})--(\ref{con:pf2:stopInDest}) thus yield $f^{st}_{ij}+f^{st}_{ji}\leq 1$, and
$F^t_{ij}+F^s_{ji}-f^{st}_{ij}-f^{st}_{ji}\in \{0,1\}$ follows.
Inequalities (\ref{con:pf2:stronger}) in combination with (\ref{con:pf2:startInSource})--(\ref{con:pf2:stopInDest}) also imply $f^{st}_{ij}=F^t_{ij}F^s_{ji}$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                     %
%         Algorithm 1                 %
%                                     %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{algorithm}
$Q\leftarrow\emptyset$\;
\Do {$\Delta Q\neq\emptyset$}{
	$(X,y,\pi)\leftarrow\text{optimal solution to $\text{LP}(\mathcal{X}_2+Q)$}$\;
	\For{$\{s,t\}\in E(D)\setminus Q$} {
		$v_{st}\leftarrow \text{optimal objective function value in $\mathcal{MF}$}$\;
	}
	$Q'\leftarrow\left\{\{s,t\}\in E(D)\setminus Q: v_{st}<1\right\}$\;
	Select some $\Delta Q\subseteq Q'$ such that $\Delta Q\neq\emptyset$ if $Q'\neq\emptyset$\;
	$Q\leftarrow Q\cup\Delta Q$\;
}
 \caption{Constraint generation}
\label{alg:cg}
\end{algorithm}

@@

Balancing the needs for rapid growth in $z\left(\text{LP}(\mathcal{X}_2+Q)\right)$ and small value of $|Q|$,
we pursue the CG scheme as follows:
In the graph $(D,Q')$, where edges correspond to pairs $\{s,t\}$ at which $v_{st}<1$, let the edge weights be $1-v_{st}$.
Let $\Delta Q$ be a matching of maximum weight in $(D,Q')$.
Thus, for each $\{u,w\}\in E(D)\setminus Q$ where (\ref{con:mf:flowNormal})--(\ref{con:vi:sumFImpSumY}) are not satisfiable given the current values of $X$ and $\pi$,
$Q$ is extended by an adjacent edge.

>>>>

Let $A(D)=A\cap(D\times D)$ be the set of ordered pairs of distinct destination nodes.
For a tree $T$ solving Problem \ref{def:problem}, define the variables $x\in\{0,1\}^{A\times A(D)}$, where
$x_{ij}^{st}=1$ iff $(i,j) \in A$ is an arc on the path from $s$ to $t$ in arborescence $T^s$.
% \newline\newline
% $x_{ij}^{st}=
% \begin{cases}
%     1 & \text{if $(i,j) \in A$ is an arc on the path from $s$ to $t$ in $T$, $(s,t)\in A(D)$},\\
%     0 & \text{otherwise}.
% \end{cases}$
% \newline\newline
Model $\mathcal{X}_2$ is strengthened by constraints on $x$, which yields model $\mathcal{X}_3$: 
\newline
\newline
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                     % 
%             MODEL X3                %
%             (4 idx vars)            %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{subequations}[resume]\label{eq:smt-x3}
\begin{align}
\notag\makebox[0pt][l]{$\displaystyle{} \min \sum\limits_{(i,j) \in A} p_{ij} \sum\limits_{s \in D} \pi^s_{ij}, $}\\
\text{s.t.} \notag \\
(\ref{con:dd:arrowFromDest}) - (\ref{con:dd:vardim}) &\text{ and }(\ref{con:dd:yvar}) - (\ref{con:vi:sumYImpSumX}), \notag\\
\label{con:mf:flowNormal} \sum\limits_{\substack{ j\in V_i}}x^{st}_{ij}-\sum\limits_{\substack{j\in V_i }}x^{st}_{ji} & = 0 & (s,t)\in A(D), i\in V\setminus\{s,t\},\\
\label{con:mf:flowDest} \sum\limits_{\substack{j\in V_t}}x^{st}_{jt}    & = 1  & (s,t)\in A(D),\\
\label{con:mf:fcap} x^{st}_{ij} &\leq X^{s}_{ij}& (i,j)\in A, (s,t)\in A(D),\\
\label{con:mf:fsym} x^{st}_{ij} &=  x^{ts}_{ji}& (i,j)\in A, (s,t)\in A(D),\\ 
\label{con:vi:sumFImpSumY} \sum_{k\in W_{ij}}x^{st}_{ik} & \leq \sum_{k\in W_{ij}} \pi^{s}_{jk} & (i,j)\in A, (s,t)\in A(D),\\
\label{con:mf:xydim} x&\in\{0,1\}^{A\times A(D)}. 
\end{align}~
\end{subequations}

Variable $x_{ij}^{st}$ is also interpreted as the \emph{flow of commodity $(s,t)$} along arc $(i,j)$.
Correspondingly, (\ref{con:mf:flowNormal})--(\ref{con:mf:flowDest}) can be viewed as flow conservation constraints,
stating that there exists a path from $s$ to $t$ in $T^s$.
Next, constraints (\ref{con:mf:fcap}) express the obvious relation between $X$ and $x$, that $(i,j)$ is an arc on the path from $s$ to $t$ only if it exists in $T^s$.
The flow symmetry equations (\ref{con:mf:fsym}) state that the path from $s$ to $t$ contains $(i,j)$ iff the reverse path contains the reverse arc $(j,i)$.
Finally, constraints (\ref{con:vi:sumFImpSumY}) reflect the fact that if the path from $s$ to $t$ intersects $(i,j)$,
then the most expensive arc leaving $i$ in $T^s$ has cost $p_{ij}$ or more.

@@

Similarly to the extension $\mathcal{X}_3$ of $\mathcal{X}_2$ by $x$-variables, the model $\mathcal{F}_2$ is extended by variables with four node indices.
Analogously to $A(D)$, let $E(D_0)=E\cap (D_0\times D_0)$ be the set of \emph{unordered} pairs of destination nodes other than $s_0$.
Define variables $f\in\{0,1\}^{A\times E(D)}$, where
$f_{ij}^{st}=f_{ji}^{ts}$ takes the value 1 iff the path in $T^0$ from $s_0$ to $s$ intersects with the path from $s_0$ to $t$ at arc $(i,j)$.
% \newline\newline
% $f^{st}_{ij}=f^{ts}_{ij}=
% \begin{cases}
%     1 & \text{if the paths in $T^0$ from $s_0$ to respectively $s$ and $t$ intersect} \\
%       & \text{at $(i,j)$,} \\
%     0 & \text{otherwise.}
% \end{cases}$
% \newline\newline
%
Variables $f_{ij}^{st}$ can also be regarded as the \emph{joint flow} from $s_0$ to $s$ and $t$ \citep{Polzin}.
In the cited work, they are successfully used to strengthen formulations for the \textsc{Minimum Steiner Tree} problem.
The new variables enable formulation of an extended model denoted $\mathcal{F}_3$:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                     %
%             MODEL F3                %
%                                     %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{subequations}[resume]
\begin{align}
\notag\min & \sum\limits_{(i,j) \in A} p_{ij} \sum\limits_{s \in D} \pi^s_{ij},\\ \notag
\text{s.t.}& && \\
\notag(\ref{con:pf1:flow})-(\ref{con:vi:sumYImpSumXTrans}) & ~\text{and} ~(\ref{con:vi:Y1}), \\
\label{con:pf2:flowHook} \sum\limits_{\substack{ j \in V_i }}\left(f^{st}_{ij}-f^{st}_{ji}\right) & \leq 
  \begin{cases}
    1, & i = s_0, \\
    0, & i\in V_0,
  \end{cases} \quad \{s,t\}\in E(D_0), \\
\label{con:pf2:startInSource}  f^{st}_{ij} & \leq F^s_{ij} \quad (i,j)\in A, \{s,t\}\in E(D_0),\\
\label{con:pf2:stopInDest}  f^{st}_{ij} & \leq F^t_{ij} \quad (i,j)\in A, \{s,t\}\in E(D_0),\\
\label{con:pf2:stronger}  F^{s}_{ij}+F^{t}_{ij}-f^{st}_{ij} &\leq g_{ij} \quad (i,j)\in A, \{s,t\}\in E(D_0),\\ 
\label{con:pf2:sumFImpSumY} \sum\limits_{k\in W_{ij}}\left(F^{t}_{ik} + F^{s}_{ki} - f^{st}_{ik} - f^{st}_{ki}\right) & \leq \sum\limits_{k\in W_{ij}}  \pi^{s}_{ik} \quad (i,j)\in A,\{s,t\}\in E(D_0),\\
\label{con:pf2:sumFImpSumY2} \sum\limits_{k\in W_{ij}}F^{t}_{ik} & \leq \sum\limits_{k\in W_{ij}}  \pi^{0}_{ik} \quad (i,j)\in A, t\in D_0,\\
\label{con:pf2:dim} f&\in\{0,1\}^{A\times E(D_0)}.
\end{align}~
\end{subequations}

By (\ref{con:pf2:flowHook}), it is ensured that the joint flow is non-increasing by increasing distance from the root $s_0$.
While constraints (\ref{con:pf2:startInSource})--(\ref{con:pf2:stopInDest}) ensure $f_{ij}^{st}=0$ if $F_{ij}^s=0$ or $F_{ij}^t=0$,
inequalities (\ref{con:pf2:stronger}) impose $f_{ij}^{st}=1$ if $F_{ij}^s=F_{ij}^t=1$.
By virtue of (\ref{con:pf2:stopInDest}), constraints (\ref{con:pf2:stronger}) replace the weaker capacity constraints (\ref{con:pf1:xfrel}).
Finally, (\ref{con:pf2:sumFImpSumY})--(\ref{con:pf2:sumFImpSumY2}) are valid inequalities justified by:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                     %
%         Proposition  		      %
%   (transformation x_{ij}^{st})      %
%                                     %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{prop}\label{prop:transX}
Assume $(F,f,g,\pi)$ satisfies (\ref{con:pf1:flow})--(\ref{con:vi:sumYImpSumXTrans}), (\ref{con:vi:Y1}), (\ref{con:pf2:flowHook})--(\ref{con:pf2:stronger}), and (\ref{con:pf2:dim}).
Let $T^0\subseteq G_g$ be the corresponding arborescence spanning $D$ (see Lemma \ref{lem:modelcorrect}),
and let $T^s$ be the arborescence obtained by redirecting the arcs in $T^0$ such that they point away from $s\in D$. 
Then, for all nodes $i$ and $j$ ($i\neq j$) spanned by $T^0$,
$$
F^t_{ij}+F^s_{ji}-f^{st}_{ij}-f^{st}_{ji} = 
	\begin{cases}
		1, ~\text{if $(i,j)$ is an arc on the path from~$s$ to $t$ in $T^s$,} \\
		0, ~\text{otherwise.}
	\end{cases}
$$
\end{prop}
%
\begin{proof}
Because $g_{ij}+g_{ji}\leq 1$, we have $F^t_{ij}+F^s_{ji}\leq 1$.
Constraints (\ref{con:pf2:startInSource})--(\ref{con:pf2:stopInDest}) thus yield $f^{st}_{ij}+f^{st}_{ji}\leq 1$, and
$F^t_{ij}+F^s_{ji}-f^{st}_{ij}-f^{st}_{ji}\in \{0,1\}$ follows.
Inequalities (\ref{con:pf2:stronger}) in combination with (\ref{con:pf2:startInSource})--(\ref{con:pf2:stopInDest}) also imply $f^{st}_{ij}=F^t_{ij}F^s_{ji}$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                     %
%         Algorithm 1                 %
%                                     %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{algorithm}
$Q\leftarrow\emptyset$\\
\Do {$\Delta Q\neq\emptyset$}{
	$(X,y,\pi)\leftarrow\text{optimal solution to $\text{LP}(\mathcal{X}_2+Q)$}$\\
	\For{$\{s,t\}\in E(D)\setminus Q$} {
		$v_{st}\leftarrow \text{optimal objective function value in $\mathcal{MF}$}$\\
	}
	$Q'\leftarrow\left\{\{s,t\}\in E(D)\setminus Q: v_{st}<1\right\}$\\
	Select some $\Delta Q\subseteq Q'$ such that $\Delta Q\neq\emptyset$ if $Q'\neq\emptyset$\\
	$Q\leftarrow Q\cup\Delta Q$\\
}
 \caption{Constraint generation}
\label{alg:cg}
\end{algorithm}

@@

Balancing the needs for rapid growth in $z\left(\text{LP}(\mathcal{X}_2+Q)\right)$ and small value of $|Q|$,
we pursue the CG scheme as follows:
In the graph $(D,Q')$, where $Q'=\{\{s,t\}\in E(D):v_{st}<1\}$,
%edges correspond to pairs $\{s,t\}$ at which $v_{st}<1$, 
let the edge weights be $1-v_{st}$.
Let $\Delta Q$ be a matching of maximum weight in $(D,Q')$.
Thus, for each $\{u,w\}\in E(D)\setminus Q$ where constraints (\ref{con:mf:flowNormal})--(\ref{con:vi:sumFImpSumY}) are not satisfiable given the current values of $X$ and $\pi$,
$Q$ is extended by an adjacent edge.

### 2018-12-27 smt/sec7-heur.tex
MI: Remarks from minor revision incorporated - mostly typos and grammar.

\textcolor{blue}{\section{Computing near-optimal solutions} \label{sec:heur}}

\textcolor{blue}{
Instances of size that prohibits computation of the optimal solution should be approached by some fast, but possibly inexact method.
Imposing a time bound on an IP-solver, and applying it to any of the models introduced in Section \ref{sec:ILP}, readily gives such a method.
Alternatively, a heuristic method dedicated to SMT can be developed.}

\textcolor{blue}{
To the best of our knowledge, the scientific literature offers few heuristic methods for SMT besides those studied by \citet{ivanova16isco}.
However, recent progress by \citet{pajor18} on heuristic methods for the related minimum Steiner tree problem can smoothly be exploited in order to generate analogous methods for SMT.
The main algorithmic components in the most basic heuristic are a construction algorithm, a local improvement algorithm,
and an algorithm for merging two Steiner trees in order to create a better one.
Construction and local improvement are applied in order to generate Steiner trees of presumably low cost.
If the cost is sufficiently low, the tree is stored in an ordered set $\mathcal{T}$ of at most $\gamma$ distinct trees.
Diversity among the trees is achieved by perturbing the edge costs randomly before construction and local improvement are applied.
After each tree generation, the new tree is merged with a randomly chosen tree from $\mathcal{T}$, 
and the resulting tree is kept in $\mathcal{T}$ if its cost is smaller than some other tree already in $\mathcal{T}$.
If also $|\mathcal{T}|=\gamma$, the most expensive tree $T_{\gamma}$ is removed from $\mathcal{T}$.
}

\textcolor{blue}{
Adapting the method by \citet{pajor18} to our problem is straightforward,
given the recently published construction and local improvement algorithms for SMT \citep{ivanova16isco}.
Let \textit{SMTHeur}$(G,D,p)$ denote the output produced when $(G,D,p)$ is input to the construction algorithm,
and the constructed tree is input to the local improvement method.
The adapted method is then described as follows, where $\mathcal{T}=\left(T_1,\ldots,T_{|\mathcal{T}|}\right)$
and $c\left(T_1\right)\leq\dots\leq c\left(T_{|\mathcal{T}|}\right)$:}

\begin{algorithm}
\textcolor{blue}{
\While{the time assigned to the heuristic is not expired} {
  \textbf{for} $\{i,j\}\in E$ \textbf{do} $\hat{p}_{ij}\leftarrow\text{random number uniformly distributed on
  $\left[\frac{1}{2}p_{ij},\frac{3}{2}p_{ij}\right]$}$\;
  $T\leftarrow\text{\textit{SMTHeur}$(G,D,\hat{p})$}$\;
  \textbf{if} $|\mathcal{T}|<\gamma$ \textbf{then} $\mathcal{T}\leftarrow \mathcal{T}\cup \{T\}$\;
  \Else {
    \textbf{if} $c(T)<c(T_{\gamma})$ \textbf{then} $\mathcal{T}\leftarrow \mathcal{T}\setminus\{T_{\gamma}\}$, $\mathcal{T}\leftarrow \mathcal{T}\cup \{T\}$\;
    $r\leftarrow\text{random number uniformly distributed on $\{1,\ldots,\gamma\}$}$\;
    \textbf{for} $\{i,j\}\in E_T\cap E_{T_r}$ \textbf{do} $\tilde{p}_{ij}\leftarrow p_{ij}$\;
    \For{$\{i,j\}\in \left(E_T\cup E_{T_r}\right)\setminus \left(E_T\cap E_{T_r}\right)$}{
      $\nu\leftarrow\text{random number uniformly distributed on $\{100,\ldots,500\}$}$\;
      $\tilde{p}_{ij}\leftarrow \nu p_{ij}$\;
    }
    \textbf{for} $\{i,j\}\in E\setminus E_T\setminus E_{T_r}$ \textbf{do} $\tilde{p}_{ij}\leftarrow 1000p_{ij}$\;
    $\tilde{T}\leftarrow\text{\textit{SMTHeur}$(G,D,\tilde{p})$}$\;
    \textbf{if} $c(\tilde{T})<c(T_{\gamma})$ \textbf{then} $\mathcal{T}\leftarrow \mathcal{T}\setminus\{T_{\gamma}\}$, $\mathcal{T}\leftarrow \mathcal{T}\cup \{\tilde{T}\}$\;
  }
}
\Return $T_1$
}

\textcolor{blue}{
\caption{Outline of the heuristic method}
\label{alg:meta}
}
\end{algorithm}

\textcolor{blue}{
Following \citet{pajor18}, the perturbed cost $\hat{p}_{ij}$ of edge $\{i,j\}$ is put equal to a random number with expected value $p_{ij}$.
Further, in the merge process, the adjusted weights $\tilde{p}_{ij}$ are determined in order to 
promote edges that appear in both $T$ and the randomly selected tree $T_{\gamma}$ in $\mathcal{T}$. 
The cost of edges appearing in exactly one of the trees are thus magnified.
For the purpose of diversification, the factor by which the cost is multiplied is chosen randomly.
To discourage the appearance of edges outside both $T$ and $T_r$ in the merge $\tilde{T}$,
their cost is magnified by a factor dominating the factors drawn for edges that occur in one of the trees.
}

\textcolor{blue}{
A few minor simplification over the approach taken by \citet{pajor18} are made in Alg.\ \ref{alg:meta}:
Rather than selecting the tree to leave $\mathcal{T}$ randomly, we consistently remove $T_{\gamma}$ when a new tree is added.
Also, for each new tree $T$ constructed, Alg.\ \ref{alg:meta} makes only one attempt to merge it with some $T_r$ stored in $\mathcal{T}$.
This contrasts the method in the cited work, where $\tilde{T}$ is chosen as the best tree found in a sequence of merges.
} \\

>>>>

\section{Computing near-optimal solutions} \label{sec:heur}

Instances of size that prohibits computation of the optimal solution should be approached by some fast, but possibly inexact method.
Imposing a time bound on an IP-solver, and applying it to any of the models introduced in Section \ref{sec:ILP}, readily gives such a method.
Alternatively, a heuristic method dedicated to SMT can be developed.

To the best of our knowledge, the scientific literature offers few heuristic methods for SMT besides those studied by \citet{ivanova16isco}.
However, recent progress by \citet{pajor18} on heuristic methods for the related minimum Steiner tree problem can smoothly be exploited in order to generate analogous methods for SMT.
The main algorithmic components in the most basic heuristic are a construction algorithm, a local improvement algorithm,
and an algorithm for merging two Steiner trees in order to create a better one.
Construction and local improvement are applied in order to generate Steiner trees of presumably low cost.
If the cost is sufficiently low, the tree is stored in an ordered set $\mathcal{T}$ of at most $\gamma$ distinct trees.
Diversity among the trees is achieved by perturbing the edge costs randomly before construction and local improvement are applied.
After each tree generation, the new tree is merged with a randomly chosen tree from $\mathcal{T}$, 
and the resulting tree is kept in $\mathcal{T}$ if its cost is smaller than some other tree already in $\mathcal{T}$.
If also $|\mathcal{T}|=\gamma$, the most expensive tree $T_{\gamma}$ is removed from $\mathcal{T}$.

Adapting the method by \citet{pajor18} to our problem is straightforward,
given the recently published construction and local improvement algorithms for SMT \citep{ivanova16isco}.
Let \textit{SMTHeur}$(G,D,p)$ denote the output produced when $(G,D,p)$ is input to the construction algorithm,
and the constructed tree is input to the local improvement method.
The adapted method is then described as follows, where $\mathcal{T}=\left(T_1,\ldots,T_{|\mathcal{T}|}\right)$
and $c\left(T_1\right)\leq\dots\leq c\left(T_{|\mathcal{T}|}\right)$:

\begin{algorithm}
\While{the time assigned to the heuristic is not expired} {
  \textbf{for} $\{i,j\}\in E$ \textbf{do} $\hat{p}_{ij}\leftarrow\text{random number uniformly distributed on
  $\left[\frac{1}{2}p_{ij},\frac{3}{2}p_{ij}\right]$}$\\
  $T\leftarrow\text{\textit{SMTHeur}$(G,D,\hat{p})$}$\\
  \textbf{if} $|\mathcal{T}|<\gamma$ \textbf{then} $\mathcal{T}\leftarrow \mathcal{T}\cup \{T\}$\\
  \Else {
    \textbf{if} $c(T)<c(T_{\gamma})$ \textbf{then} $\mathcal{T}\leftarrow \mathcal{T}\setminus\{T_{\gamma}\}$, $\mathcal{T}\leftarrow \mathcal{T}\cup \{T\}$\\
    $r\leftarrow\text{random number uniformly distributed on $\{1,\ldots,\gamma\}$}$\\
    \textbf{for} $\{i,j\}\in E_T\cap E_{T_r}$ \textbf{do} $\tilde{p}_{ij}\leftarrow p_{ij}$\\
    \For{$\{i,j\}\in \left(E_T\cup E_{T_r}\right)\setminus \left(E_T\cap E_{T_r}\right)$}{
      $\nu\leftarrow\text{random number uniformly distributed on $\{100,\ldots,500\}$}$\\
      $\tilde{p}_{ij}\leftarrow \nu p_{ij}$\\
    }
    \textbf{for} $\{i,j\}\in E\setminus E_T\setminus E_{T_r}$ \textbf{do} $\tilde{p}_{ij}\leftarrow 1000p_{ij}$\\
    $\tilde{T}\leftarrow\text{\textit{SMTHeur}$(G,D,\tilde{p})$}$\\
    \textbf{if} $c(\tilde{T})<c(T_{\gamma})$ \textbf{then} $\mathcal{T}\leftarrow \mathcal{T}\setminus\{T_{\gamma}\}$, $\mathcal{T}\leftarrow \mathcal{T}\cup \{\tilde{T}\}$\\
  }
}
\Return $T_1$

\caption{Outline of the heuristic method}
\label{alg:meta}
\end{algorithm}

Following \citet{pajor18}, the perturbed cost $\hat{p}_{ij}$ of edge $\{i,j\}$ is put equal to a random number with expected value $p_{ij}$.
Further, in the merge process, the adjusted weights $\tilde{p}_{ij}$ are determined in order to 
promote edges that appear in both $T$ and the randomly selected tree $T_{\gamma}$ in $\mathcal{T}$. 
The cost of edges appearing in exactly one of the trees are thus magnified.
For the purpose of diversification, the factor by which the cost is multiplied is chosen randomly.
To discourage the appearance of edges outside both $T$ and $T_r$ in the merge $\tilde{T}$,
their cost is magnified by a factor dominating the factors drawn for edges that occur in one of the trees.

A few minor simplification over the approach taken by \citet{pajor18} are made in Alg.\ \ref{alg:meta}:
Rather than selecting the tree to leave $\mathcal{T}$ randomly, we consistently remove $T_{\gamma}$ when a new tree is added.
Also, for each new tree $T$ constructed, Alg.\ \ref{alg:meta} makes only one attempt to merge it with some $T_r$ stored in $\mathcal{T}$.
This contrasts the method in the cited work, where $\tilde{T}$ is chosen as the best tree found in a sequence of merges.
